### 2018-05-09 morusAC.tex
citations, common.sty

\newif\ifanonymous
%\anonymoustrue%uncomment for submission version

\newif\ifshrink\shrinkfalse
%\shrinktrue for screen-friendly original LNCS paper size

\documentclass[a4paper]{llncs}
\usepackage{common}

\ifanonymous
\author{}
\institute{}
\else
\author{Tomer Ashur \and Maria Eichlseder \and Ga\"etan Leurent \and Brice Minaud \and Yann Rotella \and Yu Sasaki \and Beno\^it Viguier}
\institute{}
%\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
\fi

\begin{document}

\title{\morus: Biases in the stream}
\maketitle

\input{morusAC_00_Abstract}
\input{morusAC_01_Introduction}
\input{morusAC_02_Preliminaries}
\input{morusAC_03_MiniMorus}
\input{morusAC_04_Mini_trails}
\input{morusAC_05_Full_trails}
\input{morusAC_06_Discussion}
\input{morusAC_07_IniFin}

\ifanonymous
\else
  %\subsubsection{Acknowledgments.}
  %...
  %\horizonack{The research leading to these results has received funding from
  %the European Union's Horizon 2020 research and innovation programme under
  %grant agreement No 644052 (HECTOR) and \dots.}
\fi

\bibliographystyle{splncs03}
\bibliography{morusAC}

\appendix

\input{morusAC_App_IniFin}

\end{document}