### 2018-06-21 slides/tikz/trail-alpha.tex
Benoit Lunch Talk

\begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
  \rotwfalse
  \messagefalse
  \printstate
  \draw[trail, alpha]
    (C) -- node[right] {$i$} (lll-1)
    (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
    (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
    (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
    (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
    (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
    ;
  \draw[trail, alpha, dashed]
    (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east);
  % \draw[alpha, very thick] (W40) + (0,-0.2) ellipse (0.6cm and 0.175cm);

  \draw (W42) + (0,-0.5) node[below] {weight($\alpha^t_i$) = 1 (not 2)};

\end{tikzpicture}%}}}