### 2018-09-06 AC2018_camera_ready/morusAC.tex
Camera-ready version

\newif\ifanonymous
\newif\iflncsmargins
%\anonymoustrue%uncomment for submission version

\newif\ifshrink
%\shrinktrue % for screen-friendly original LNCS paper size

\documentclass[a4paper]{llncs}
\usepackage{common}
\pagestyle{plain}
\usepackage{multirow}

\ifanonymous
\lncsmarginstrue
\author{}
\institute{}
\else
\author{Tomer Ashur\inst{1} \and Maria Eichlseder\inst{2} \and Martin M. Lauridsen \and Ga\"etan Leurent\inst{3} \and Brice Minaud\inst{4} \and Yann Rotella\inst{3} \and Yu Sasaki\inst{5} \and Beno\^it Viguier\inst{6}}
\institute{
imec-COSIC, KU Leuven, Belgium
\and
Graz University of Technology, Austria
\and
Inria, France
\and
Royal Holloway University of London, United Kingdom
\and
NTT, Japan
\and
Radboud University, Netherlands\\
\path|tomer.ashur@esat.kuleuven.be|, \path|maria.eichlseder@iaik.tugraz.at|, \path|mail@martinlauridsen.info|, \path|gaetan.leurent@inria.fr|, \path|brice.minaud@gmail.com|, \path|yann.rotella@inria.fr|, \path|b.viguier@science.ru.nl|, \path|sasaki.yu@lab.ntt.co.jp|
}
\fi

\iflncsmargins
\else
%\usepackage[hmargin=2.5cm,vmargin=3cm]{geometry}
\fi

\pagestyle{empty}

\begin{document}

\title{Cryptanalysis of \MORUS}
\maketitle
%\thispagestyle{empty}

\input{morusAC_00_Abstract}
\input{morusAC_01_Introduction}
\input{morusAC_02_Preliminaries}
\input{morusAC_03_MiniMorus}
\input{morusAC_04_Mini_trails}
\input{morusAC_05_Full_trails}
\input{morusAC_06_Discussion}
\input{morusAC_07_IniFin}
\input{morusAC_08_Conclusion}

\bibliographystyle{splncs03}
\bibliography{morusAC}

\ifanonymous
\section*{Auxiliary Material}
\else
%\clearpage
\fi

\appendix

\input{morusAC_App_IniFin}

\end{document}