### 2018-11-30 slides-asiacrypt/tikz/VerticalMorus.tex
finalize slides

\vspace{0.1cm}
\begin{tikzpicture}[xscale=1.2,yscale=1.2]
  \pgfmathsetmacro{\roundsep}{0.80}

  \foreach \w in {-1,...,5} {
    \foreach \r in {-1,...,5} {
      \coordinate (W\r\w) at (\w,-\r*\roundsep-.125);
    }
    \coordinate (W-2\w) at (\w,--2.375*\roundsep-.425);
  }
  \foreach \w in {0,...,4} {
    %\draw (W-2\w) -- (W4\w);
    \node[inner sep=1pt] (S\w) at ($(W0\w) +(0,.1+1.7)$) {$S_\w$};
  }
  \draw (W-1-1) node[inner sep=1pt] {\phantom{$M$}};
  \draw (W-1-1) node[above right,gray,box,minimum height=4.8ex,yshift=1.1ex,xshift=-.5ex] (filt) {\Filt};
  \foreach \w in {0,...,3} {
    \draw (W-2\w|-filt.south) +(0,.85ex+.7*\w ex) coordinate[tee, name=t\w];
    \draw[next] (t\w) -- (t\w-|filt.east);
  }
  \draw (W22) node[gray,box,minimum width=5cm,minimum height=5cm,yshift=3ex] (upd) {\StateUpdate};
  \foreach \w in {0,...,4} {
    \draw[next] (W-2\w) -- (W-2\w|-upd.north);
    \draw[next] (W5\w|-upd.south) -- (W5\w);
  }
  \draw[next] (filt.south) -- ++(0,-2ex) node[below] {out};
  \draw[prev] (upd.west) -- (upd-|filt) -- ++(0,2ex) node[above] {in};
\end{tikzpicture}